Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-01-10 manuscript/manuscript.Rmd
Set seed for reproducibility.
Prevent printing code.
Adapted figure labels and inserted figure caption into chunk.

library(power.transform)

@@

@@

Power transformation does not guarantee that transformed variables are normally distributed. In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure \@ref(fig:reduced-normality).

``` {r reduced-normality, echo="FALSE", include="FALSE", fig.cap="Power transformations may reduce normality depending on location. Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. Panel A and B show the $\\lambda$ value found through maximum likelihood estimation for Box-Cox and Yeo-Johnson transformations, respectively."}

@@

p <- ggplot2::ggplot(
  data = data,
  mapping = ggplot2::aes(
    x = mu,
    y = lambda,
    colour = method))
p <- p + ggplot2::theme_light(base_size=base_size)
p <- p + ggplot2::geom_point()
p <- p + ggplot2::geom_hline(yintercept = 1.0, linetype="longdash", colour="grey40")
p <- p + ggplot2::annotate(
  geom = "text",
  x = 6,
  y = 1,
  label = "expected",
  colour=annotation_settings$colour,
  family=annotation_settings$family,
  fontface=annotation_settings$face,
  size=annotation_settings$geom_text_size,
  vjust=1.5,
  hjust=1.0)
p <- p + ggplot2::scale_x_continuous(labels = scales::math_format())

@@

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

>>>>

# Allow for defining figure captions within a chunk.
knitr::opts_knit$set(
  eval.after = "fig.cap")

require(power.transform)
require(ggplot2)
require(data.table)
require(latex2exp)

@@

# Set seed.
set.seed(19L)

@@

Power transformation does not guarantee that transformed variables are normally distributed. In fact, depending on location and scale of the variable, power transformation may reduce normality, as shown in Figure.

``` {r reduced-normality, echo=FALSE, fig.cap=cap}
cap <- paste0(
  "Power transformations may reduce normality depending on location. ",
  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
  "and even breaks down due to numerical issues for $\\mu > 10^3$.")

@@

p <- ggplot2::ggplot(
  data = data,
  mapping = ggplot2::aes(
    x = mu,
    y = lambda,
    colour = method))
p <- p + ggplot2::theme_light(base_size=base_size)
p <- p + ggplot2::geom_point()
p <- p + ggplot2::geom_hline(yintercept = 1.0, linetype="longdash", colour="grey40")
p <- p + ggplot2::annotate(
  geom = "text",
  x = 6,
  y = 1,
  label = "expected",
  colour=annotation_settings$colour,
  family=annotation_settings$family,
  fontface=annotation_settings$face,
  size=annotation_settings$geom_text_size,
  vjust=1.5,
  hjust=1.0)
p <- p + ggplot2::scale_x_continuous(
  name = latex2exp::TeX("$\\mu$"),
  labels = scales::math_format())

p <- p + ggplot2::scale_y_continuous(
  name = latex2exp::TeX("$\\lambda$"))

@@
